The fifteen year struggle of decentralizing privacy-enhancing technology

### 2014-03-25 paper.tex
tabel inline en other changes

% Title of the paper
\title{The fifteen year struggle of decentralizing privacy-enhancing technology}
% Authors
\author{Rolf Jagerman, Wendo Sab\'ee, Laurens Versluis, Martijn de Vos}
\date{}

@@

\section{Introduction}
	All feelings of privacy concerning browsing the internet, talking on the telephone or location tracking of cellphones are an illusion. In recent years the need for privacy-enhancing technology has become more apparent. Revelations by Edward Snowden of government misconduct and constitutional violations have sent shockwaves through the internet.

@@

	The most widely used anonymous communication system is Tor. In this paper we will analyze Tor and its semi-centralized nature. Tor struggles to keep up with the bandwidth demands of its users. As the number of users increases, the need to decentralize Tor becomes more urgent. Decentralizing Tor isn't an easy task: After 15 years of decentralization attempts, the network is still centralized. Only few decentralized alternatives to Tor exist, however they lack the userbase to be considered safe and useful. Examples include Gnutella \cite{ripeanu2001peer}, Freenet \cite{lua2005survey} and Tapestry.

@@

	In 2002, the Tor project discontinued their old code base and reimplement the project as Tor, the Second Generation Onion Router. They introduced perfect forward secrecy, directory servers, hidden services and more \cite{dingledine2004tor}. In this section, we will explain the various components of Tor, the structure of the network, circuit creation and disadvantages of Tor.

@@

	\subsection{Onion routing}
		As described in the original design paper of The Onion Router \cite{goldschlag1996hiding}, network traffic is forwarded through a circuit of nodes, where each node only knows the previous and next node in the circuit. With a sufficiently long circuit of (independent) nodes, this means that two communicating parties can remain oblivious of each other's physical location.

		Say we have a circuit consisting of four nodes: our trusted client node ($C$), an (entry or guard) relay node ($X$), a (middle) relay node ($Y$) and an (exit) relay node ($Z$). In this case, there are three relay nodes, but this is not necessarily always the case as more middle relay nodes can be added. A visualisation of this path can be seen in figure \ref{fig:tor_layout}. Each node has its own public key and a corresponding private key. When building the circuit, our client generates a distinct secret for each of these nodes. More information about the circuit setup, can be found in \ref{ss:tor_circuit}.

@@

		\subsection{Tor stinks?}
			Tor is not only used by human rights activists. It is also used by distributors of illegal content and providers of illegal services, because it is deemed untraceable. This serious problem is disrupting Tor's public image. A recent example is the shutdown of \emph{Silk Road}, which is an online market for trading prohibited substances. The market operated as a hidden Tor service. These services are accessed through an onion address and not an IP-address, hiding the physical location of the service. This makes it very hard for agencies to track and shut down these operations.
			% Meer voorbeelden:
			% - international business times -> silk road (zie wikipedia)
			% - weapons of mass destruction (?)

@@

	\subsection{Incentives in decentralized systems}
		Tax evasion and environmental pollusion can be seen as forms of free-riding, a phenomenon that is prominent in Tor. People predominantly use more Tor bandwidth than they donate, see section \ref{ss:tor_disadvantages}. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor, all have failed. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words, the reputation data has to be accurate and reliable. Besides the integrity of this data, the traffic it generates on the network should have minimal impact on the overall performance.

@@

	\subsection{Bootstrapping new nodes}
		If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be a starting point where this new node can discover neighbors in the network to connect with. In Tor, a directory server can tell the new node what his neighbors are and where to find them \cite{dingledine2004tor}.

		Moving this system to a peer-to-peer base is difficult: Dingledine et al. stated that this is indeed still an open problem. With decentralized systems, there is no central directory server to tell a new node where to locate neighbors. Some systems such as Tarzan, MorphMix and Pastry \cite{rowstron2001pastry, rennhard2002introducing} are decentralized but they do suffer from performance issues.

@@

\section{Decentralized privacy-enhancing systems}
	\label{sec:decentralized}
	Fully decentralized systems with large scale usage are without exception based on the peer-to-peer paradigm. Many such systems have been proposed, yet only some have been implemented and are currently in use and actively maintained \cite{mislove2004ap3, rennhard2002introducing, panchenko2006nisan, rowstron2001pastry, nambiar2006salsa, freedman2002tarzan, ripeanu2001peer, androutsellis2004survey}. Here we focus on fully decentralized networks with the exception of Torsk (which is almost fully decentralized but still requires a Neighbourhood Authority).

@@

@@

	\subsection{AP3}
		AP3 (Anonymizing Peer-to-Peer Proxy) \cite{mislove2004ap3} makes cooperative, decentralized anonymous communication possible. The AP3 system provides clients with three primitives: anonymous message delivery, anonymous channels and secure pseudonyms. Users are able to send and receive unicast, multicast and anycast messages anonymously. The strategy that AP3 is using for message delivery is similar to that of Tarzan: it relies on a network of peers to forward messages. A node along the request path, does not know whether the node from which it receives a message is the message's originator or simply another forwarding peer.

	In this section, we discussed existing decentralized peer-to-peer networks. In this section we will make a comparison between these networks, as seen in table \ref{tbl:comparison}. For each network, we will explain the following features:

@@

\section{The documented struggle of alternative internet projects}
	\label{sec:comparison}
	The largest repository of decentralization attempts is located at http://redecentralize.org. The struggle and pains of these projects illustrates the difficulty of decentralization. No projects succeeded in creating an alternative internet infrastructure. This ranges from self hosted cloud applications, to crypto currencies, to anonymous networks.

@@

@@

	% Tabel kan inline tabel worden, zonder "last activity" en woordje "commit(s)"
	\begin{table}[p]
		\centering
		\include{tbl-alternativeinternet}
		\caption{A list with the top 10 projects from the Alternative Internet repository as of 2014-03-15, sorted on the number of commits.}
		\label{tbl:altinternet}
	\end{table}
\end{landscape}

>>>>

% Title of the paper
\title{The fifteen year struggle of decentralizing privacy-enhancing technology}
% Authors
\author{Rolf Jagerman, Wendo Sab\'ee, Laurens Versluis, Martijn de Vos\\ Supervised by Dr. Ir. Johan Pouwelse}
\date{}

@@

\section{Introduction}
	All feelings of privacy concerning browsing the internet, talking on the telephone or location tracking of cellphones are an illusion. In recent years the need for privacy-enhancing technology has become more apparent. Revelations by Edward Snowden of government misconduct and constitutional violations have sent shock waves through the internet.

@@

	The most widely used anonymous communication system is Tor. In this paper we will analyze Tor and its semi-centralized nature. Tor struggles to keep up with the bandwidth demands of its users. As the number of users increases, the need to decentralize Tor becomes more urgent. Decentralizing Tor isn't an easy task: After 15 years of decentralization attempts, the network is still centralized. Only few decentralized alternatives to Tor exist, however they lack the user base to be considered safe and useful. Examples include Gnutella \cite{ripeanu2001peer}, Freenet \cite{lua2005survey} and Tapestry.

@@

	In 2002, the Tor project discontinued their old code base and re-implement the project as Tor, the Second Generation Onion Router. They introduced perfect forward secrecy, directory servers, hidden services and more \cite{dingledine2004tor}. In this section, we will explain the various components of Tor, the structure of the network, circuit creation and disadvantages of Tor.

@@

	\subsection{Onion routing}
		As described in the original design paper of The Onion Router \cite{goldschlag1996hiding}, network traffic is forwarded through a circuit of nodes, where each node only knows the previous and next node in the circuit. With a sufficiently long circuit of (independent) nodes, this means that two communicating parties can remain oblivious of each others physical location.

		Say we have a circuit consisting of four nodes: our trusted client node ($C$), an (entry or guard) relay node ($X$), a (middle) relay node ($Y$) and an (exit) relay node ($Z$). In this case, there are three relay nodes, but this is not necessarily always the case as more middle relay nodes can be added. A visualization of this path can be seen in figure \ref{fig:tor_layout}. Each node has its own public key and a corresponding private key. When building the circuit, our client generates a distinct secret for each of these nodes. More information about the circuit setup, can be found in \ref{ss:tor_circuit}.

@@

		\subsection{Tor stinks?}
			Tor is not only used by human rights activists. It is also used by distributors of illegal content and providers of illegal services, because it is deemed untraceable. This serious problem is disrupting Tor's public image. A recent example is the shutdown of \emph{Silk Road}, which is an online market for trading prohibited substances, buying weapons of mass destruction and other illegal goods. The market operated as a hidden Tor service. These services are accessed through an onion address and not an IP-address, hiding the physical location of the service. This makes it very hard for agencies to track and shut down these operations.
			% Meer voorbeelden:
			% - international business times -> silk road (zie wikipedia)
			% - weapons of mass destruction (?)
			% Is de zin nu zo goed? Ik heb wat voorbeelden erbij gezet

@@

	\subsection{Incentives in decentralized systems}
		Tax evasion and environmental pollution can be seen as forms of free-riding, a phenomenon that is prominent in Tor. People predominantly use more Tor bandwidth than they donate, see section \ref{ss:tor_disadvantages}. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor, all have failed. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words, the reputation data has to be accurate and reliable. Besides the integrity of this data, the traffic it generates on the network should have minimal impact on the overall performance.

@@

	\subsection{Bootstrapping new nodes}
		If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be a starting point where this new node can discover neighbours in the network to connect with. In Tor, a directory server can tell the new node what his neighbours are and where to find them \cite{dingledine2004tor}.

		Moving this system to a peer-to-peer base is difficult: Dingledine et al. stated that this is indeed still an open problem. With decentralized systems, there is no central directory server to tell a new node where to locate neighbours. Some systems such as Tarzan, MorphMix and Pastry \cite{rowstron2001pastry, rennhard2002introducing} are decentralized but they do suffer from performance issues.

@@

\section{Decentralized privacy-enhancing systems}
	\label{sec:decentralized}
	Fully decentralized systems with large scale usage are without exception based on the peer-to-peer paradigm. Many such systems have been proposed, yet only some have been implemented and are currently in use and actively maintained \cite{mislove2004ap3, rennhard2002introducing, panchenko2006nisan, rowstron2001pastry, nambiar2006salsa, freedman2002tarzan, ripeanu2001peer, androutsellis2004survey}. Here we focus on fully decentralized networks with the exception of Torsk (which is almost fully decentralized but still requires a Neighborhood Authority).

@@

			\subsection{AP3} %2004
		AP3 (Anonymizing Peer-to-Peer Proxy) \cite{mislove2004ap3} makes cooperative, decentralized anonymous communication possible. The AP3 system provides clients with three primitives: anonymous message delivery, anonymous channels and secure pseudonyms. Users are able to send and receive unicast, multicast and anycast messages anonymously. The strategy that AP3 is using for message delivery is similar to that of Tarzan: it relies on a network of peers to forward messages. A node along the request path, does not know whether the node from which it receives a message is the message's originator or simply another forwarding peer.

@@

	\subsection{Comparison}
	In this subsection we will make a comparison between these networks, as seen in table \ref{tbl:comparison}. For each network, we will explain the following features:

@@

	\end{landscape}

\section{The documented struggle of alternative internet projects}
	\label{sec:comparison}

	The largest repository of decentralization attempts is located at http://redecentralize.org. The struggle and pains of these projects illustrates the difficulty of decentralization. No projects succeeded in creating an alternative internet infrastructure. This ranges from self hosted cloud applications, to crypto currencies, to anonymous networks.

@@

			% Tabel moet nog mooi gemaakt worden
	\begin{table}
		\centering
		\include{tbl-alternativeinternet}
		\caption{A list with the top 10 projects from the Alternative Internet repository as of 2014-03-15, sorted on the number of commits.}
		\label{tbl:altinternet}
	\end{table}

	\newpage

@@

### 2014-03-25 tbl-alternativeinternet.tex
tabel inline en other changes

\begin{tabular}[p]{| r | l | p{3cm} | p{3cm} | p{3cm} | p{3cm} | p{3cm} | p{3cm} |}
	\hline
	\textbf{\#} & \textbf{Name} & \textbf{Language} & \textbf{Age} & \textbf{Last activity} & \textbf{LOC} & \textbf{Commits} & \textbf{Contributors} \\ \hline

	1 & ownCloud & PHP & 6 year(s) & 2014-03-15 & 1,297 K & 34,391 commit(s) & 392 contributor(s) \\ \hline
	2 & Freenet & Java & 13 year(s) & 2014-03-15 & 442 K & 32,009 commit(s) & 183 contributor(s) \\ \hline
	3 & Tor & C & 12 year(s) & 2014-03-13 & 329 K & 29,200 commit(s) & 184 contributor(s) \\ \hline
	4 & GNUnet & C & 8 year(s) & 2014-03-13 & 427 K & 21,398 commit(s) & 37 contributor(s) \\ \hline
	5 & StatusNet & PHP & 6 year(s) & 2014-03-14 & 230 K & 14,877 commit(s) & 92 contributor(s) \\ \hline
	6 & Diaspora* & Ruby & 3 year(s) & 2014-03-14 & 51 K & 14,202 commit(s) & 368 contributor(s) \\ \hline
	7 & SlapOS & Python & 8 year(s) & 2014-03-15 & 588 K & 14,046 commit(s) & 93 contributor(s) \\ \hline
	8 &Tahoe-LAFS & Python & 7 year(s) & 2014-03-13 & 158 K & 11,565 commit(s) & 54 contributor(s) \\ \hline
	9 & Tribler & Python & 8 year(s) & 2014-03-15 & 148 K & 11,521 commit(s) & 48 contributor(s) \\ \hline
	10 & Lorea & PHP & 6 year(s) & 2014-03-12 & 873 K & 11,066 commit(s) & 96 contributor(s) \\ \hline
\end{tabular}

>>>>

\begin{tabular}[p]{| r | l | p{2.2cm} | p{1.2cm} | p{2.2cm} | p{1.2cm} | p{1.5cm} | p{2.3cm} |}
	\hline
	\textbf{\#} & \textbf{Name} & \textbf{Language} & \textbf{Age} & \textbf{Last activity} & \textbf{LOC} & \textbf{Commits} & \textbf{Contributors} \\ \hline

	1 & ownCloud & PHP & 6 year(s) & 2014-03-15 & 1,297 K & 34,391 & 392 \\ \hline
	2 & Freenet & Java & 13 year(s) & 2014-03-15 & 442 K & 32,009 & 183 \\ \hline
	3 & Tor & C & 12 year(s) & 2014-03-13 & 329 K & 29,200 & 184 \\ \hline
	4 & GNUnet & C & 8 year(s) & 2014-03-13 & 427 K & 21,398 & 37 \\ \hline
	5 & StatusNet & PHP & 6 year(s) & 2014-03-14 & 230 K & 14,877 & 92 \\ \hline
	6 & Diaspora* & Ruby & 3 year(s) & 2014-03-14 & 51 K & 14,202 & 368 \\ \hline
	7 & SlapOS & Python & 8 year(s) & 2014-03-15 & 588 K & 14,046 & 93 \\ \hline
	8 &Tahoe-LAFS & Python & 7 year(s) & 2014-03-13 & 158 K & 11,565 & 54 \\ \hline
	9 & Tribler & Python & 8 year(s) & 2014-03-15 & 148 K & 11,521 & 48 \\ \hline
	10 & Lorea & PHP & 6 year(s) & 2014-03-12 & 873 K & 11,066 & 96 \\ \hline
\end{tabular}
